---
title: "Correlations"
author: "BPR"
date: "2023-03-24"
categories: [R]
---

This example is really just so I can practice putting R chunks into my blog and making sure everything works as I expect it to. I might even remove this blog after I'm better at this.

The mythical "everyone" knows to run a correlation, and to look at a scatterplot, as part of your exploratory data analysis (EDA). Amost all of these correlations are Pearson correlations, which are great for linear relationships, but not so great for nonlinear relationships. However, a distance correlation can start to get at the nonlinear. Here's an example.

```{r createData}
set.seed(20230324)
runif(n = 50,
      min = 1,
      max = 20) -> # Yes, I write my R code
  x                #  backwards. Deal with it.
1.7 * (x-10) ^ 2 -2 + rnorm(n = 50,
                            mean = 0,
                            sd = 2) ->
  y
```

Quadratics are perhaps the "simplest" nonlinear relationship.

```{r savePlot, include = FALSE}
jpeg(filename = "image.jpg")
plot(x,y,
     pch = 16,
     cex = 0.5,
     main = "Nonlinear relationship")
dev.off()
```

```{r plotData, include = FALSE}
plot(x,y,
     pch = 16,
     cex = 0.5,
     main = "Nonlinear relationship")
```

!\[\](quadratic.jpg)

Yep, looks nonlinear.

```{r correlations, include = FALSE}
library(energy)     # Distance correlation
cor(x, y)
dcor(x, y)
```

The Pearson correlation, `r cor(x,y)` , fails to indicate a relationship but the distance correlation (energy::dcor) `r dcor(x, y)` does indicate a relationship. Bootstrapped confidence intervals would show that there is significance in the distance correlation value.

Wow, that wasn't bad; render this document, then render the website homepage, and it works! Of course, there is no thumbnail. I'll have to figure out how to use the \[image\] tab to have the plot be the thumbnail. (I could, as a workaround, save the plot, and then insert the saved plot. However, that seems clunky and open to problems. That's for another day.)
